---
title: "MSCA 31008 HW5"
author: "Ruoxin Shi"
date: "2022/3/5"
output: html_document
---

## Cluster-wise Regression - GermanCredit

### Load Data

Load data from the library.

```{r GermanCredit}
# load data
library(caret)
data(GermanCredit)
```

Use only the numeric independent variables 1, and 3 through 7 as predictors.

```{r data, echo=TRUE}
set.seed(1123)
# drop unnecessary variables
# selected only numerical variables as predictors

# Standardize your variables
scaled_data <- scale(GermanCredit[,c(1:7)])

# Use the 6 numeric variables other than "amount" as independent variables.
cleaned_df <- scaled_data[,c(1, 3:7)]
Amount <- scaled_data[,2]

# final data
final_data <- as.data.frame(cbind(Amount, cleaned_df))

# split
train_ind <- sample(seq_len(nrow(final_data)), size = 632)
train <- final_data[train_ind, ]
test <- final_data[-train_ind, ]
```

### Import Function

```{r}
# import the function
source("clustreg.txt")
source("clustreg.predict.txt")

```

### Create Model

Perform cluster-wise regression for 1, 2, and 3 clusters. 

```{r model}
clustreg_credit_1 <- clustreg(train,1,1,1123,1)
clustreg_credit_2 <- clustreg(train,2,2,1123,10)
clustreg_credit_3 <- clustreg(train,3,2,1123,10)
training_r2 <- c(clustreg_credit_1$rsq.best,clustreg_credit_2$rsq.best,clustreg_credit_3$rsq.best)
```


### Holdout Validation

Using the 3 models built with 1, 2, and 3 clusters respectively, predict classes for the holdout set.

```{r predict}
pred1 <- clustreg.predict(clustreg_credit_1, test)
pred2 <- clustreg.predict(clustreg_credit_2, test)
pred3 <- clustreg.predict(clustreg_credit_3, test)
holdout_r2 <- c(pred1$rsq, pred2$rsq, pred3$rsq)

```

Show a table of training R^2, holdout R^2, and percentage decrease from train to holdout R squared: (Train R square – Holdout R square)/ Train R square.

```{r}
pct_decrease <- (training_r2 - holdout_r2)/training_r2
result <- data.frame(training_r2, holdout_r2, pct_decrease)
result
```


Plot the R^2 for different number of clusters.

```{r plot}
plot(c(1,2,3), training_r2,
     ylim=c(0,1),type="l",col=4,main="VAF Plot for GermanCredit Data: Cluster-wise Regression",
     ylab="Variance Accounted For",xlab="Number of Clusters")
lines(c(1,2,3), holdout_r2, col="green")
legend("topleft", pch="-", col=c(4, "green"), c("train", "holdout"), cex=1)

```


**Which model is the most stable? Which performs the best on the holdout set. Which one would you choose as the final model?**

According to the percent decrease from train to holdout R^2, the most stable model is the one with 3 cluster. Also, this is the one that performs best on the holdout set. So I would choose this as the final model.


### Clustering Result

```{r}
cluster2 <- clustreg_credit_2$cluster
cluster3 <- clustreg_credit_3$cluster
table(cluster2, cluster3)
```
```{r}
prop.table(table(cluster2, cluster3))
```
From the above result, we could see that there are some overlap in the first cluster 
but little overlap in the second cluster. But we could not simply say that the second and third clusters in the 
3-cluster model are splitted by the second cluster in the 2-cluster model. They are still different clusters.




```{r result1}
clustreg_credit_2$results

```


```{r result2}
clustreg_credit_3$results

```


### Summarize the result


**From the training plot of R squared, which solution seems best?**\
Based on the plot, the solution with 3 clusters works best with the highest VAF value.\


**How did each solution perform in holdout?**\
Similarly, in holdout data, the solution with 3 clusters performs better.\

**Are you able to interpret the results by reading the regression coefficients?**\
**Can you tell what types of clusters have been formed?**\
\
2 cluster result for training set:\

1. In the first cluster, the amount starts from a low value. The most important variable here are *"Duration"* and *"InstallmentRatePercentage"*, with the prior having positive effect and the later having postive effect on *"Amount"*. *"Age"* is also important, with a slightly positive effect on *"Amount"*.\

2. In the second cluster, the amount starts from a much higher value. The most important variable here is still *"Duration"*, which has a much greater positive effect on *"Amount"* than the previous cluster. *"InstallmentRatePercentage"* is also important, with a higher level of negative effect on *"Amount"* as well. *"Age"* is also important here, with a slightly higher positive effect on *"Amount"*. The difference between this cluster and the previous one is that in this cluster, *"NumberExistingCredits"* plays a role as well, with a negative effect.\


3 cluster result for training set:\

1. In the first cluster, the amount starts from a very low value. Similar to the 2-cluster result, the most important variables are *"Duration"* and *"InstallmentRatePercentage"*, with the prior having positive effect and the later having postive effect on *"Amount"*. *"ResidenceDuration"* is also important, with a slightly positive effect on *"Amount"*. The rest variables are less important in this cluster.\

2. In the second cluster, the amount starts from a low value, but higher than the previou cluster. The most important variable is *'Duration'*, which has a positive effect on the *'Amount'*. All other variables plays some roles in this cluster. However, the effect of *"InstallmentRatePercentage"* decreases a lot and becomes the least important one among all others.\

3. In the third cluster, the amount starts from a much higher value. The most important variable is still *'Duration'*, which has a positive effect on the *'Amount'*. *"InstallmentRatePercentage"* and *"ResidenceDuration"* also plays important roles. *"Age"* and *"NumberExistingCredits"* are less important but still have a slighter degree of impact. \



So from the above interpretation we could know that in different clusters, same variable might have various influence on the target variable. But among all variables, *'Duration'* always plays a significant role.
The amount increases as duration increases in both the 2-cluster and 3-cluster result. This means that the general influence of *"Duration"* is similar for different type of clusters. The difference is in the level of impact.\


Also, in the 2-cluster result, we could hardly see the influence of *"ResidenceDuration"*, *"NumberExistingCredits"* and *"NumberPeopleMaintenance"*. But in the 3-cluster result, it shows clearly that *"ResidenceDuration"* has a positive effect on *"Amount"* in the first and third cluster, and a negative effect in the second cluster; both *"NumberExistingCredits"* and *"NumberPeopleMaintenance"* are more important in the second cluster with both positive effect, and not very important in the first and third cluster. This situation might also indicated that 3-cluster model is better since it clarifies the impact of variables in a clearer and more comprehensive way.\
\

